---
title: "REMIND validation"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
params:
  mif: ""
  cfg: ""
  extraColors: true
  warning: false
  message: false
---

```{r include=FALSE}
library(piamValidation)
library(knitr)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(plotly)
library(gridExtra)

knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  message = params$message,
  warning = params$warning
)
```

## Import and Prepare Data

Loading data from:

```{r}
for (m in params$mif) cat(paste0(normalizePath(m), "\n"))
```

Using config:

```{r}
cat(params$cfg, "\n")
```


```{r, message = FALSE, warning = TRUE}
# import scenario and reference data once, it is used for the validation and
# for the line plots; importScenarioData() removes the "+" notation from
# variable names so that they match the variable names in the validation results
data <- piamValidation:::importScenarioData(params$mif)

valiData <- validateScenarios(data, params$cfg, extraColors = params$extraColors)
valiData <- appendTooltips(valiData)

# add "<model>_smoothed" reference data in which the 2020 value is replaced by
# the 2018-2022 average to dampen the Covid shock; only used in the line plots
hist_smoothed <- filter(data, scenario == "historical") %>%
  piamValidation:::average_2020() %>%
  filter(grepl("_smoothed$", model), !is.na(value))
data <- bind_rows(data, hist_smoothed)
```

# Validation

```{r}
# heatmap wrapper: report missing validation results instead of failing
REMIND_heatmap <- function(var, vD = valiData) {
  d <- filter(vD, variable == var)
  if (nrow(d) == 0) {
    cat("No validation results for", var, "\n")
    return(invisible(NULL))
  }
  validationHeatmap(d)
}

# lineplot wrapper: one plot per region, reg = "all" plots every region in
# which the variable was validated, regions without results are skipped
REMIND_lineplot <- function(var,
                            reg = "World",
                            vD = valiData,
                            df = data,
                            hist_models = NULL,
                            xlimits = c(2000, 2035)) {

  vD <- filter(vD, variable == var)
  if (nrow(vD) == 0) {
    cat("No validation results for", var, "\n")
    return(invisible(NULL))
  }

  scen_data <- filter(df, scenario != "historical", variable == var)
  hist_data <- filter(df, scenario == "historical", variable == var)
  if (!is.null(hist_models)) {
    hist_data <- filter(hist_data, model %in% hist_models)
  }

  # regions with validation results for this variable, World first
  if (identical(reg, "all")) {
    regions <- unique(as.character(vD$region))
    if (is.factor(vD$region)) regions <- intersect(levels(vD$region), regions)
  } else {
    regions <- reg
  }

  plot_list <- htmltools::tagList()
  for (r in regions) {
    line_data <- filter(vD, region == r)
    if (nrow(line_data) == 0) {
      cat("No validation results for", var, "in region", r, "\n")
      next
    }
    plot_list[[length(plot_list) + 1]] <- linePlotThresholds(
      line_data,
      refData  = filter(hist_data, region == r),
      scenData = filter(scen_data, region == r),
      xlim = xlimits)
  }
  plot_list
}
```

## Emissions

### Transport {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Emi|CO2|w/o Bunkers|Energy|Demand|Transport"
REMIND_heatmap(var)
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("CEDS", "EDGARghg", "CEDS_smoothed", "EDGARghg_smoothed"))
```

### Buildings {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Emi|CO2|Energy|Demand|Buildings"
REMIND_heatmap(var)
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("CEDS", "EDGARghg", "CEDS_smoothed", "EDGARghg_smoothed"))
```

### Energy and IP {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Emi|CO2|w/o Bunkers|Energy and Industrial Processes"
REMIND_heatmap(var)
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("CEDS", "EDGARghg", "CEDS_smoothed", "EDGARghg_smoothed"))
```

### Energy Supply {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Emi|CO2|Energy|Supply"
REMIND_heatmap(var)
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("CEDS", "EDGARghg", "CEDS_smoothed", "EDGARghg_smoothed"))
```



## PE

### Coal {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "PE|Coal"
REMIND_heatmap(var)
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", 
                hist_models = c("IEA-EB-directSum", "BP", "IEA-EB-directSum_smoothed", "BP_smoothed"))
```

### Gas {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "PE|Gas"
REMIND_heatmap(var)
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", 
                hist_models = c("IEA-EB-directSum", "BP", "IEA-EB-directSum_smoothed", "BP_smoothed"))
```

### Oil {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "PE|Oil"
REMIND_heatmap(var)
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", 
                hist_models = c("IEA-EB-directSum", "BP", "IEA-EB-directSum_smoothed", "BP_smoothed"))
```


## SE 

### Electricity {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "SE|Electricity"
REMIND_heatmap(var)
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("Ember", "BP", "Ember_smoothed", "BP_smoothed"))
```

## FE 

### Total {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "FE"
REMIND_heatmap(var)
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("IEA-EB-directSum", "IEA-EB-directSum_smoothed"))
```


## Capacities

### Coal {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Cap|Electricity|Coal"
REMIND_heatmap(var)
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("Ember", "Ember_smoothed"))
```

### Gas {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Cap|Electricity|Gas"
REMIND_heatmap(var)
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("Ember", "Ember_smoothed"))
```

### Solar {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Cap|Electricity|Solar"
REMIND_heatmap(var)
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("Ember", "Ember_smoothed"))
```

### Wind {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Cap|Electricity|Wind"
REMIND_heatmap(var)
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("Ember", "Ember_smoothed"))
```

### Hydro {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Cap|Electricity|Hydro"
REMIND_heatmap(var)
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("Ember", "IRENA", "Ember_smoothed", "IRENA_smoothed"))
```

### Nuclear {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Cap|Electricity|Nuclear"
REMIND_heatmap(var)
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("Ember", "Ember_smoothed"))
```

## EVs

### Stocks {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Stock|Transport|Pass|Road|LDV|Four Wheelers|BEV"
REMIND_heatmap(var)
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all")
```

### Sales {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Sales|Transport|Pass|Road|LDV|Four Wheelers|BEV"
REMIND_heatmap(var)
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all")
```

## Carbon Management

### Storage {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Carbon Management|Storage"
REMIND_heatmap(var)
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all")
```


# Sustainability Concerns

```{r, fig.width=7, fig.height=6}
var <- "PE|Biomass"
REMIND_heatmap(var)
```

## Summary {.tabset}

```{r}
df_summary <- valiData

colors <- c(green     = "#008450",
            yellow    = "#EFB700",
            red       = "#B81D13",
            cyan      = "#66ccee",
            blue      = "#4477aa",
            grey      = "#808080")

# find "critical == yes" data points of each color
summary <- filter(df_summary, region == "World") %>%
  dplyr::count(model, check) %>%
  group_by(model) %>%
  mutate(percent = n / sum(n) * 100)

# change order of colors
summary$check <- factor(
  summary$check, 
  levels = rev(c("green", "yellow", "cyan", "red", "blue", "grey"))
  )

# stacking order and midpoints
summary <- summary %>%
  arrange(model, desc(check)) %>%
  group_by(model) %>%
  mutate(
    cumulative = cumsum(n),
    midpoint = cumulative - (n / 2)
  )


# stacked bar plot in absolute numbers of checks
p1 <- ggplot(summary, aes(x = model, y = n, fill = check)) +
  geom_bar(stat = "identity") + # Absolute stacked bars
  scale_fill_manual(values = colors) + # Custom colors
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 0.5)) +
  geom_text(aes(label = n, y = midpoint), color = "white", size = 3.5) +
  labs(y = "Count", x = "Model") +
  ggtitle("World") + 
  theme(legend.position = "none")

# normalized
p2 <- ggplot(summary, aes(x = model, y = n, fill = check)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_fill_manual(values = colors) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5)) +
  geom_text(aes(label = paste0(round(percent, 1), "%"),
                group = check),
            position = position_fill(vjust = 0.5),
            color = "white",
            size = 3.5) +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Percentage", x = "Model") +
  ggtitle("World") + 
  theme(legend.position = "none")


# find "critical == yes" data points of each color
summary <- filter(df_summary, region != "World") %>%
  dplyr::count(model, check) %>%
  group_by(model) %>%
  mutate(percent = n / sum(n) * 100)

# change order of colors
summary$check <- factor(
  summary$check, 
  levels = rev(c("green", "yellow", "cyan", "red", "blue", "grey"))
  )

# stacking order and midpoints
summary <- summary %>%
  arrange(model, desc(check)) %>%
  group_by(model) %>%
  mutate(
    cumulative = cumsum(n),
    midpoint = cumulative - (n / 2)
  )

# stacked bar plot in absolute numbers of checks
p3 <- ggplot(summary, aes(x = model, y = n, fill = check)) +
  geom_bar(stat = "identity") + # Absolute stacked bars
  scale_fill_manual(values = colors) + # Custom colors
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 0.5)) +
  geom_text(aes(label = n, y = midpoint), color = "white", size = 3.5) +
  labs(y = "Count", x = "Model") +
  ggtitle("Regions") + 
  theme(legend.position = "none")

# normalized
p4 <- ggplot(summary, aes(x = model, y = n, fill = check)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_fill_manual(values = colors) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5)) +
  geom_text(aes(label = paste0(round(percent, 1), "%"),
                group = check),
            position = position_fill(vjust = 0.5),
            color = "white",
            size = 3.5) +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Percentage", x = "Model") +
  ggtitle("Regions") + 
  theme(legend.position = "none")
```


### Percentage

```{r, fig.width=7, fig.height=4}
grid.arrange(p2, p4, nrow = 1)
```

### Count
```{r, fig.width=7, fig.height=4}
grid.arrange(p1, p3, nrow = 1)
```
